---
title: "Homework 2"
author: "Fahim Hussain"
date: "September 8, 2018"
output:
  pdf_document: default
  html_document: default
---
```{r warning = FALSE, message= FALSE}
library(MASS)
library(ggplot2)
library(dplyr)
library(cowplot)
```

```{r}
head(Boston, 5)
```
```{r}
dim(Boston)
```

####How many rows are in this data set? How many columns? What do the rows and columns represent?
####There are 506 rows and 14 columns. The rows represent the observations and the columns represents the variables in a data set. For example, row 1 represents all the observed values from different variables(columns).

#b) Make some pairwise scatterplots of the predictors (columns) in this data set. Describe your findings.
```{r message=FALSE, fig.width= 10, fig.height = 7}
Boston %>% select(crim, dis, black,lstat,medv) %>% pairs()
```

####I chose these 5 variables, crim, dis, black,lstat, and medv because I found them to be most interesting. I believed they would have some sort of relationship with each other. Some observations we can make is that, 1) crime rates are higher when the mean distance to the employment centers are smaller. This is probably because the area is most likely an urban area, which should have a higher crime rate than the rural areas. We can also see that the 2) lower status of the population decreases as the distance to the employment centers increase. This is probably due to the fact that lower status population are offered more work near employment centers. Consequently, 3) As the median value of occupied homes increases, crime rates decrease, distance to employment centers increase, and lower status of population decreases. This can most likely be attributed to the fact that wealthier individuals will tend to purchase more expensive homes in a more rural area, where crime rates is not significant. 4) The relationship between black population and the other variables are not apparent, there does not seem to be much significant relationship.

```{r message=FALSE, fig.width= 10, fig.height = 7}
Boston %>% select(nox, age, ptratio, indus, zn) %>% pairs()
```

####I chose to examine these 5 variables because I would like a general idea of the area in general. Immediately, 1) we can see that as the proportion of owner-occupied units built prior to 1940(age) increases, the concentration of nitrogen oxides(nox) increases. It isn't a surprise that older homes will contain a lot of bacteria and harmful substances that can have a big effect on the air. Surprisingly, 2) we see that as the ptratio increases, the proportion of non-retail business acres per town(indus) increases. This could be possibly be explained by a larger population due to a higher ratio of pupils to students. 3) There does not seem to be any other significant analysis from the other graphs provided. 

#c)
```{r}
cor(Boston)[,1]
```



####Are any of the predictors associated with per capita crime rate? If so, explain the relationship.
####Yes, as we have seen, there are many predictors associated with per capita crime rate. We can see that crimes are more prevalent near employment centers. Also, houses with a lower mean average value tend to be surrounded by a higher crime rate area. We can see some of these relationship from the above table. Obviously, crime will have a one-to-one relationship with itself. Crime has the biggest relationship with tax and rad. This is probably because high property tax rate will mean an expensive house and accebility to radial highways will allow criminals an easy escape.

#d)
```{r message=FALSE}
Boston %>% ggplot(aes(x=crim))+geom_histogram(fill="Red")
summary(Boston$crim)
```

#### Do any of the suburbs of Boston appear to have particularly high crime rates? Tax rates? Pupil-teacher ratios? Comment on the range of each predictor. 

#### Although most of the crime rates are near 0 or close to it, we see some neighborhoods with an abnormally high crime rate. From the summary statistics, the median crime rate is about 0.26% but the mean is 3.61%, which means we have a positive skew in the distribution.This is seen by the graph, and the skew heavily influence the mean. We can inference that there are certain suburbs with an abnormally high crime rate. Crime rate is bounded below by the minimum of 0.00632 and bounded above by the maximum of 88.9762. The high differential is a clear indicator that we have an abnormally high amount of crime rate in certain suburbs.

```{r message=FALSE}
plot_grid(Boston %>% ggplot(aes(x=tax))+geom_histogram(fill="Blue")+labs(title="Property Tax rate"),
          Boston %>% ggplot(aes(x=ptratio))+geom_histogram(fill="Green")+
            labs(title="Pupil-Teacher Ratio"),labels="AUTO")
summary(Boston$tax)
summary(Boston$ptratio)
```

#### The median and mean are nearly identical for the ptratio and the tax is almost similar. However, we do see some suburbs have a high tax rate than normal and a high ptratio than normal. The tax range is from (187,711), which is a huge range. This is most likely due to the income difference among the individuals and the property values that they contain. The ptratio ranges from (12.6,22), which is not much of a difference. A difference of 10 is a fair value for the ratio.

#e)

```{r}
Boston %>% filter(chas==1) %>% nrow()
```
####  How many of the suburbs in this data set bound the Charles river? 
#### There are 35 suburbs that bound the Charles River

#f)
```{r}
median(Boston$ptratio)
```

#### What is the median pupil-teacher ratio among the towns in this data set?
#### The median is 19.05

#g)

```{r}
min(Boston$medv)
Boston[Boston$medv==5,]
summary(Boston)
```
####Which suburb of Boston has lowest median value of owneroccupied homes? What are the values of the other predictors for that suburb, and how do those values compare to the overall ranges for those predictors? Comment on your findings.

####399 and 406 are the suburbs with the lowest median value of owner occupied homes. I first looked for the minimum median value and then used that value to isolate the suburbs that contain that minimum. From the summary of Boston, we can see that 399 and 406 have an abnormally high amount of crime rate compared to average. The proportion of non-retail businesses are also significantly higher than the mean and median of the overall data. The age variable is also at its max for both 399 and 406, which means the houses are all built prior to 1940. This explains why both the suburbs have a higher nox value compared to average. However, the distance to an employment center is less than average. Despite their old, noxious homes, 399 and 406 have a significantly high tax rate compared to average. The ptratio and black values align to the average of the overall dataset. However, the lower status of the population greatly exceeds the average and median.

#h)

```{r}
Boston %>% filter(rm >7) %>% nrow()
Boston %>% filter(rm >8) 
```
#### In this data set, how many of the suburbs average more thanseven rooms per dwelling? More than eight rooms per dwelling? Comment on the suburbs that average more than eight rooms per dwelling.

#### 64 suburbs average more than 7 rooms per dwelling and 13 average more than 8. Those with more than 8 rooms per dwelling tend to have lower crime rate, the proportion of non-retail business acres tend to be on the lower side, most do not tract by the Charles River, the nox values seem to be distributed normally, most of the houses tend to be built prior to 1940, the distance from an employment center seems to be about normally distributed among the 13 suburbs, the tax seems to be lower than the average, the black population and ptratio aligns with the median, but the lower status of the population is on the low side compared to the median and average.