---
title: "Homework 1"
author: "Fahim Hussain"
date: "August 28, 2018"
output:
  html_document: default
  pdf_document: default
---
```{r echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
```

```{r}
ggplot(data=mpg)
```

**1) There is nothing to see when I run ggplot(data=mpg). I need to provide a visualization to ggplot to show the dta.**

```{r}
nrow(mpg)
ncol(mpg)
```
**2) There are 234 rows and 11 columns in the mpg dataset.**

```{r}
?mpg
```

**3) The drv variable is a categorical value that tells you whether the car is a front-wheel drive, f, a rear wheel drive, r, or a 4 wheel drive, 4.**

```{r}
ggplot(data=mpg, aes(x=cyl, y=hwy))+ geom_point()
```

**4) Graph above**

```{r}
ggplot(data=mpg, aes(x=class, y=drv)) + geom_point()
```

**5) This plot is not useful because there is no significant analysis in examining the relationship of a car's class and its drv. It also does not utilize our entire dataset efficiently because there are only 12 data points in the chart.**  
